\chapter{INTRODUCTION}

Since the start of the scientific thinking, mathematics has been of the key
languages in which we can understand the world, and, Partial Differential
Equations (PDEs) are one of the main dialects we use to model and to simplify
nature's complexity.

These equations do not always have simple closed-form solutions (that is, when
they have), showing the need of approximated numerical solutions coming from
numerical methods for PDEs in order to extract predictions, compare with data,
build simulations and so on. These classical methods, widely studied for all
types of equations, rely on discretizing the PDE's domain into regular grids
and performing time steps.

With the ascension of Machine Learning (ML) methods and the whole ecosystem of
tools built around the field, a natural extesion is to build a bridge
connecting numerical solution of PDEs with ML models, a field called Scientific
Machine Learning (SciML).

In this work, we aim to study the application of SciML methods for
approximating the solution of the Boussinesq System of PDEs, a shallow water
model for a water tank with small depth, coming from the Euler Equations model
of Hydrodynamics. Our goal is to implement and evaluate the performance of two
famous models of SciML: the Physics-Informed Neural Networks and the Fourier
Neural Operators, which together represent the two main types of methods in
SciML, investigating the impact of data and the PDE model (so called
Physics-Informed) presence in training, revealing their advantages and
limitations for each method.

\pagebreak

\section{Preliminaries}

Let us present, in this section, the mathematical definition for each method we
will study, as well as the Boussinesq System, which is our benchmark model of
choice.

\subsection{Multilayer Perceptrons}

In this work, we focus on Multilayer Perceptrons (MLPs), popularized by
\textcite{rumelhart1986}, which are, in essence, affine transformations
parametrized by tunable weights and biases, followed by nonlinear
transformations, in which an optimization process acts to fit a dataset of
values from the supposed target function.

Given a labeled dataset ${\{(x_i, u_i)\}}_{i=1}^{N_{data}}$, the objective is
to fit the function $u_{\theta}(x)$ by minimizing a loss function, typically
the Mean Squared Error (MSE). With $||.||_2$ denoting the Euclidian norm, we
have the unconstrained optimization problem:
\begin{align}
    \text{MSE}(\theta) & = \frac{1}{N_{data}} \sum_{i=1}^{N_{data}} {|| u_{\theta}(x_i) - u_i ||}_2^2 \label{eq:data_loss_nn}
\end{align}
\noindent where $u_{\theta}$ is the Multilayer Perceptron (MLP) is a function $u_\theta: \mathbb{R}^{d_{in}} \to \mathbb{R}^{d_{out}}$ defined by the composition:
\begin{align}
    h^{(0)}       & = x \in \mathbb{R}^{d_{in}}, \label{eq:input}                                                                      \\
    h^{(l)}       & = \sigma \left( W^{(l)} h^{(l-1)} + b^{(l)} \right) \in \mathbb{R}^{N} , \quad l = 1, \dots, L-1 \label{eq:hidden} \\
    u_{\theta}(x) & = W^{(L)} h^{(L-1)} + b^{(L)} \in \mathbb{R}^{d_{out}}. \label{eq:output}
\end{align}
\noindent where $N, L \in \mathbb{N}$ denotes the number $N$ of neurons in each of $L$ layers, $W^{(1)} \in \mathbb{R}^{d_{in} \times N}$, $W^{(L-1)} \in \mathbb{R}^{N \times d_{out}}$ and $b^{(1)} \in \mathbb{R}^{N}$, $W^{(l)} \in \mathbb{R}^{N\times N}, b^{(l)} \in \mathbb{R}^{N} \text{ for } l = 2, ..., L - 2$ are the weight matrix and bias vector of layer $l$, respectively, and $\sigma: \mathbb{R} \to \mathbb{R}$ is a nonlinear sigmoidal activation function applied element-wise, in other words, sigmoidal meaning $\sigma(t) \to 1$ as $t \to +\infty$ and $\sigma(t) \to 0$ as $t \to -\infty$ and elemnent-wise meaning it is applied to each component of a vector. We denote $\theta = \ \{W^{(l)}, b^{(l)}\}_{l=1}^{L}$ to be all of the network parameters.

A common result and of utmost importance is that, if the activation function is
sigmoidal, then MLPs are universal approximators, meaning they can approximate
any continuous function \cite{cybenko1989}.

In Figure \ref{fig:MLP_depiction}, we show a common depiction of the neural
networks. In this case, the input vector is the pair $(x,t) \in \mathbb{R}^2$
and output is $(\eta_{\theta}(x,t), u_{\theta}(x,t)) \in \mathbb{R}^2$.
\begin{figure}[h!]
    \centering
    \begin{tikzpicture}[
            x=1cm,
            y=1cm,
            nnnode/.style={circle, draw, line width=0.8pt, minimum size=0.40cm, inner sep=0pt},
            inout/.style={nnnode, fill=black, draw=black},
            hidden/.style={nnnode, fill=gray, draw=gray},
            conn/.style={draw=gray!70, line width=0.35pt}
        ]

        % Input
        \node[inout] (x1) at (0,-2.4) {};
        \node[inout] (x2) at (0,-3.2) {};
        \node at (0,0.15) {\small Input $(x,t)$};

        % Hidden layer 1
        \foreach \i in {1,...,6} {
                \node[hidden] (h1_\i) at (2.5,-0.8*\i) {};
            }

        % Hidden layer 2
        \foreach \i in {1,...,6} {
                \node[hidden] (h2_\i) at (5.0,-0.8*\i) {};
            }
        \node at (5.0,0.15) {\small Hidden layers};

        % Hidden layer 3
        \foreach \i in {1,...,6} {
                \node[hidden] (h3_\i) at (7.5,-0.8*\i) {};
            }

        % Output
        \node[inout] (out1) at (10.0,-2.4) {};
        \node[inout] (out2) at (10.0,-3.2) {};
        \node at (10.0,0.15) {\small  Output $(\eta_{\theta}, u_{\theta})$};

        % Connections
        \foreach \j in {1,...,6} {
                \draw[conn] (x1) -- (h1_\j);
                \draw[conn] (x2) -- (h1_\j);
            }

        \foreach \i in {1,...,6} {
                \foreach \j in {1,...,6} {
                        \draw[conn] (h1_\i) -- (h2_\j);
                        \draw[conn] (h2_\i) -- (h3_\j);
                    }
            }

        \foreach \i in {1,...,6} {
                \draw[conn] (h3_\i) -- (out1);
                \draw[conn] (h3_\i) -- (out2);
            }

    \end{tikzpicture}
    \caption{Example of MLP architecture of input vector $(x, t)$, with $L=3$ hidden layers and $N=6$ neurons on each layer.}
    \label{fig:MLP_depiction}
    \autoriapropria
\end{figure}

\subsection{Physics-Informed Neural Networks}

Introduced by \textcite{raissi2017}, Physics-Informed Neural Networks (PINNs)
seeks to insert the PDE directly into the optimization process, forcing the
neural network satisfy to the model equations and approximating the PDE
solution.

Consider the general PDE Initial Value Problem (IVP) form as:
\begin{equation}\label{eq:generic_pde}
    \left\{
    \begin{aligned}
        \mathcal{D}[u](x,t) & = 0, \quad \forall (x,t) \in \Omega \times (0,T], \\
        u(x,0)              & = u_0(x), \quad x \in \Omega
    \end{aligned}
    \right.
\end{equation}

\noindent where $\mathcal{D}$ is a differential operator on $u$, $\Omega \times (0, T]$ is the PDE spatio-temporal domain and $u_0$ is the initial condition suitable for $\mathcal{D}$.

The central idea for PINNs is to incorporate (\ref{eq:generic_pde}) into the
loss function, taking place of a data loss similar to equation
\ref{eq:data_loss_nn}.

Consider the following:

\begin{equation}
    \text{MSE}_\mathcal{D} = \frac{1}{N_\mathcal{D}}\sum_{i=1}^{N_\mathcal{D}}
    \left| \mathcal{D}[u_\theta](x_i,t_i)\right|^2,
\end{equation}
\noindent where $\{(x_i,t_i)\}_{i = 1}^{N_\mathcal{D}}$ are points in the PDEs spatio-temporal domain $\Omega \times (0, T]$.

We can also force the MLP to fit the initial condition by

\begin{equation}
    \text{MSE}_0 = \frac{1}{N_0}\sum_{i=1}^{N_0} \left| u_\theta(x_i,0) - u_0(x_i) \right|^2,
\end{equation}
\noindent where $\{x_i\}_{i=1}^{N_0}$ are points in the $\Omega \times \{0\}$ domain ($T=0$).

In \textcite{raissi2017}, the points $\{(x_i,t_i)\}_{i = 1}^{N_\mathcal{D}}$
and $\{x_i\}_{i=1}^{N_0}$ are called collocation points at PDE's domain and
initial condition, respectively.

Thus, the loss function for PINNs is simply:

\begin{equation}
    \text{MSE} = \text{MSE}_\mathcal{D} + \text{MSE}_0 \label{eq:loss_pinn}
\end{equation}

Due to the compositional nature of MLPs, implementation of PINNs often use
Automatic Differentiation (AD) \parencite{linnainmaa1970, paszke2019pytorch} to
compute gradients. This technique takes advantage of Chain Rule derivatives to
build a computational graph of derivatives until reaching basic terms with
straightforward and exact derivatives.

Also, common optimizers for the PINN loss (\ref{eq:loss_pinn}) include the
Adaptive Moment Estimation (ADAM) \parencite{kingma2014adam} and the
Limited-Memory Broyden-Fletcher-Goldfarb-Shanno (L-BFGS)
\parencite{nocedal1980lbfgs}. These two algorithms were already used by
\textcite{raissi2017}.

\subsection{Fourier Neural Operators}

Presented by \textcite{li2020fourier}, Fourier Neural Operator (FNO) aims to
learn maps between function spaces, coming from Neural Operators (NOs), a
generalizaiton of NNs for approximating maps between infinite-dimensional
spaces.

The target (potentially nonlinear) operator is defined by, given $D \in
    \mathbb{R}^{d_{in}}$ a bounded, open set, $G^{\dagger} : \mathcal{A} [D,
        \mathbb{R}^{d_{a}} ] \to \mathcal{U} [D, \mathbb{R}^{d_{u}}]$, where
$\mathcal{A}$ and $\mathcal{U}$ Banach spaces, a potentially
infinite-dimensional complete vector space with respect to some norm
\parencite{kreyszig1978introductory},

We will study the example coming from a parametrized PDE with real parameters,
approximating the operator that takes a parameter value and outputs the PDE
solution to that parameter.

The Fourier Neural Operator, $G_\theta$, is an iterative architecture mimicking a Neural Network of the form equation \ref{eq:hidden}, meaning it composes linear operators followed by element-wise nonlinearities:
\begin{equation}
    G_\theta = Q \circ \sigma\Big(W_L + \mathcal{K}_L\Big) \circ \cdots \circ \sigma\Big(W_1 + \mathcal{K}_1\Big) \circ P.
\end{equation}

\noindent where, the lifting operator $P$ maps the input function $a(x,t) \in \mathcal{A} [D, \mathbb{R}^{d_{a}} ]$ to a higher-dimensional latent representation
$P(a)(x,t) = v_0(x,t) \in \mathbb{R}^{d_c}$, where $d_c$ is the dimension of the so called channels, usually way higher than the original dimension of $u$; 
the Spectral Layers, $\sigma\Big(W_l + \mathcal{K}_l\Big)$ for $l=1, \ldots, L-1$, in which $W_l \in \mathbb{R}^{d_c \times d_c} \text{ for } l=2, \ldots, L-2$ 
is a skip-conection meant for simple matrix multiplations, $\mathcal{K}_l$ is an Integral operator.

After the final layer, the projection operator $Q$ maps this representation back to $Q(v_\ell)(x,t)= u(x,t)$. Here, $\theta$ is meant to represent the many parameters of the $G_\theta$ operator, which will be defined along.

Usually, $P$ and $Q$ are parametrized by a encoding/decoding shallow neural
networks, with the goal of simply raising the input into a higher dimension.

The core of the FNO is the stack of Spectral Layers. Each layer updates the
last layer $v_\ell$ by linear transformations, followed by a non-linear
point-wise activation $\sigma$:
\begin{equation}
    v_{\ell+1}(x,t) = \sigma \Big( (W_\ell\,v_\ell)(x,t) + (\mathcal{K}_\ell v_\ell)(x,t) \Big).
\end{equation}
Here, $W_\ell$ is a point-wise operator that applies a linear transformation to the channel dimension $d_c$ at each coordinate $(x,t)$. Also, $\mathcal{K}_\ell$ is a general integral operator (linear) acting via a kernel function $\kappa_\ell$:
\begin{equation}
    (\mathcal{K}_\ell v_\ell)(x,t) = \int_{D} \kappa_\ell(x, t, y, \tau)\,v_\ell(y,\tau)\,dy\,d\tau,
\end{equation}
where $\kappa_\ell$ is the learned kernel of the operator (usually a shallow neural network) and $D$ is the domain of the problem.

The key point of FNO is to suppose that $\kappa_\ell(x, t, y,
    \tau)=\kappa_\ell(x - y, t - \tau)$. This allows the integral to be computed
efficiently through the frequency domain, since it becomes a convolution
(assuming time is periodic, which in our problem it is):
\begin{equation}
    (\mathcal{K}_\ell v_\ell)(x,t) \overset{\text{FNO}}{=} \int_{D} \kappa_\ell(x - y, t - \tau)\,v_\ell(y,\tau)\,dy\,d\tau = \mathcal{F}^{-1} \Big( \mathcal{F}(\kappa_\ell) \cdot \mathcal{F}(v_\ell) \Big)(x,t).
    \label{eq:fno_approximation}
\end{equation}

An inverse Fourier Transform then returns the result to the physical domain.
The Fourier transform used in \eqref{eq:fno_approximation} is simply the
Fourier series, which is defined in the following.

\subsection{Learning the Kernel in the Frequency Domain}
To see how FNO learns the kernel, we examine how the convolution is calculated
using the Fourier Series. For a periodic function $f(x)$ on $[0, 2\pi]$, its
series representation is:
\[
    f(x) = \sum_{k \in \mathbb{Z}} \hat{f}_k \, e^{ikx}, \quad \text{where} \quad \hat{f}_k = \frac{1}{2\pi} \int_0^{2\pi} f(x) e^{-ikx} dx.
\]
One can check that the convolution results in the point-wise product of
frequencies:
\begin{align*}
    (\kappa * v)(x) & = \int_0^{2\pi} \kappa(x-y) v(y) dy                                \\
                    & = \sum_{k \in \mathbb{Z}} (2\pi \hat{\kappa}_k \hat{v}_k) e^{ikx}.
\end{align*}
This is the central idea of FNO. Instead of learning the kernel function $\kappa_\ell$ in the physical domain, FNOs directly parametrizes its Fourier coefficients, discarding learning the whole kernel as a neural net. The learnable complex-valued weights $R_{\ell,k}$ are precisely this parametrization:
\[
    R_{\ell,k} \approx 2\pi \hat{\kappa}_{\ell,k}.
\]
In practice, computationally, this operation is performed only on a truncated
set of the most important, low-frequency modes, where $|k| \le k_{\max}$, which
is a hyperparameter.

\subsection{Loss Function}

The training objective is to minimize the error between the network's
prediction $G_\theta(c_i)$ and the observed solution $u_i$ from the dataset.
The loss function used is the relative $L^2$-norm. For functions, this is
defined on the continuous domain $D$:
\begin{equation}
    \mathcal{L}(\theta) = \frac{1}{N} \sum_{i=1}^N \frac{\| G_\theta(c_i) - u_i \|_{L^2(D)}}{\| u_i \|_{L^2(D)}}.
\end{equation}
However, in the implementation, both $G_\theta(c_i)$ and $u_i$ are discretized as matrices (grid 64x64) converted into vectors, much how is done in works with images, with the relative error calculated with the usual $L^2$-norm.